---
title: "p8105 hw1 cgm2163"
output: github_document
---

Homework #1 Data Science I

# Problem 1

First, I will upload the tidyverse, which has already been installed on my computer.

```{r}
library(tidyverse)

```


Next, I will upload a **data frame** that samples from 
a _normal distribution_:


```{r}

hw1_df = tibble(
  
  ## numeric vector: random sample of 10 from normal distribution
  
  numbers = rnorm(10),
  
  ## logical vector returning "true" if sample is greater than 0
  
  numbers_pos = numbers > 0,
 
  ## character vector of length 10
  
  emotions = c("happy", "sad", "amazing", "depressed",
               "confused", "puzzled", "thrilled", "confident", 
               "bewildered", "bummed" ),
  
  ## factor vector of length 10 with 3 "levels"
  
  emotion_class = factor (c("good", "bad", "good", "bad", "unsure",
                            "unsure", "good", "good", "unsure", 
                            "bad"))
  ##vector of factors classifying the types of emotions in the character vector
  
  )

```


Next, we can attempt to take the **mean** of each vector:

 * The mean of _numbers_ is `r mean(pull(hw1_df, numbers))`.

 * The mean of _numbers_pos_ is `r mean(pull(hw1_df, numbers_pos))`.

However, when we attempt to find the means of _emotions_ and _emotion_class_, we recieve an error.

The code chunk below provides the error: _arguement is not numeric or logical: returning NA_

**mean(pull(hw1_df, emotions))** 

_and_

**mean(pull(hw1_df, emotion_class))**

Now, we can try to convert the variables to numeric to see if we can get a mean using the following code chunk:

**mean(as.numeric(pull(hw1_df, emotions)))**

_and_

**mean(as.numeric(pull(hw1_df, emotion_class)))**


The output provides us with another error, _NAs introduced by coercion_ for **emotions**, but produces a numerical mean, 2, for **emotion class**. This is likely because while **emotions** is a string of characters that differs for each **numbers** value, **emotion_class** can be converted essentially into numerical groups of 1, 2, and 3, providing us with a mean = 2.


# Problem 2

First, I can install the "palmerpenguins" package with the code provided in the homework.

Next, I can use the additional code provided to load the "penguins" dataset.

```{r}

data("penguins", package = "palmerpenguins")

#list all variables in data set "penguins"

ls(penguins)

#list number of columns in data set "penguins

ncol(penguins)

#list number of rows in data set "penguins

nrow(penguins)

#calculate mean flipper length in mm, ignoring NA values

mean(pull(penguins, flipper_length_mm), na.rm=TRUE)


```


We can also create a scatterplot representing the association between**bill_length_mm** and **flipper_length_mm**:



```{r}

#create a scatter plot with bill length = x and flipper length = y

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point ()

#save plot as a pdf in the folder

ggsave("bill_flipper_scatter.pdf", height = 4, width = 6)


```


**End of HW1**

